NLIZE: A Perturbation-Driven Visual Interrogation Tool for Analyzing and Interpreting Natural Language Inference Models

### 2018-03-30 paper/pipeline.tex
update figure
length fit

\begin{figure}[htbp]
\centering
\vspace{-2mm}
 \includegraphics[width=1.0\linewidth]{pipelineView}
 \vspace{-6mm}
 \caption{
The pipeline view provides the visual representation of three-stage (encoder, attention, classifier) model. 
%The pipeline view, by showing the aggregated gradient distribution, informs the user about how different parts of the model respond to the optimization.
 %
In the proposed tool, we allow the model parameter to be updated to correct a prediction error. 
The optimization hyperparameters are shown in (a), and each of the stages is illustrated in (b), where the user can click on the blue ``parameter'' area to enable or disable its update (the legend is shown in (d)). In (c), we select whether we want to use the current update setting as displayed or try all the configuration combinations as the parameter update can be enabled or disabled for each stage.
 }
\label{fig:pipelineView}
\vspace{-2mm}
\end{figure}

\subsection{Pipeline View}
\label{sec:pipeline}
The Pipeline view provides a direct visual representation of the three stages (encoder, attention, classifier) of the model. In the proposed tool, we allow the model parameter to be updated (via an optimization) to correct a prediction error (\textbf{T3}). The pipeline view, by visualizing the aggregated gradient distribution, informs the user about how each stage of the model responds to the optimization.

There are many ways to update the model to correct a prediction. A simple way to is by applying standard backpropagation that overfits to the example. However, without any constraint, the update step may alter the model in unexpected ways.
% \shusen{what is the benefit of using mira?}
Instead, we adopt the same idea in \emph{margin-infused relaxed algorithm (MIRA)}~\cite{CrammerSinger2003}, where we regulate the optimization with the L2-norm of the parameter change. Therefore, in the proposed tool, we obtain the target parameters by the following:
\begin{equation}
\underset{\mathbf{W}'}{\mathrm{argmin}}( C \mathbb{J}(\mathbf{W}') + ||\mathbf{W}' - \mathbf{W}||^2)
\end{equation}
where $\mathbb{J}(\mathbf{W})$ is the loss function of the neural model, $\mathbf{W}$ is the original model's parameters taken as constant, $\mathbf{W}'$ is the parameters to update, and $C$ is the weighting term, which determines whether we intended to emphasize more on obtaining better fitting or deviating less from the original model. Due to the nonconvex nature of the neural networks, we use SGD to optimize the above combined loss function.
%
With this formation, we try to find a good approximation to the newly assigned label, while still maintaining relatively small change with respect to the original model.

>>>>

\begin{figure}[htbp]
\centering
\vspace{-2mm}
 \includegraphics[width=0.95\linewidth]{pipelineView}
 \vspace{-3mm}
 \caption{
The pipeline view provides the visual representation of three-stage (encoder, attention, classifier) model. 
%The pipeline view, by showing the aggregated gradient distribution, informs the user about how different parts of the model respond to the optimization.
 %
In the proposed tool, we allow the model parameter to be updated to correct a prediction error. 
The optimization hyperparameters are shown in (a), and each of the stages is illustrated in (b), where the user can click on the blue ``parameter'' area to enable or disable its update (the legend is shown in (d)). In (c), we select whether we want to use the current update setting as displayed or try all the configuration combinations as the parameter update can be enabled or disabled for each stage.
 }
\label{fig:pipelineView}
\vspace{-2mm}
\end{figure}

\subsection{Pipeline View}
\label{sec:pipeline}
The Pipeline view provides a direct visual representation of the three stages (encoder, attention, classifier) of the model. In the proposed tool, we allow the model parameter to be updated (via an optimization) to correct a prediction error (\textbf{T3}). The pipeline view, by visualizing the aggregated gradient distribution, informs the user about how each stage responds to the optimization.

There are many ways to update the model to correct a prediction. The simplest approach is applying standard backpropagation and overfits to the example. However, without any constraint, the update step may alter the model in unexpected ways.
% \shusen{what is the benefit of using mira?}
Instead, we adopt the idea in \emph{margin-infused relaxed algorithm (MIRA)}~\cite{CrammerSinger2003}, where we regulate the optimization with the L2-norm of the parameter change. Therefore, in the proposed tool, we obtain the target parameters by the following:
\begin{equation}
\underset{\mathbf{W}'}{\mathrm{argmin}}( C \mathbb{J}(\mathbf{W}') + ||\mathbf{W}' - \mathbf{W_o}||^2)
\end{equation}
where $\mathbb{J}(\mathbf{W})$ is the loss function of the neural model, $\mathbf{W_o}$ is the original model's parameters taken as constant, $\mathbf{W}'$ is the parameters to update, and $C$ is the weighting term, which determines whether we intended to emphasize more on obtaining better fitting or deviating less from the original model. Due to the nonconvex nature of the neural networks, we use SGD to optimize the above combined loss function.
%
With this formulation, we try to find a good approximation to the newly assigned label, while still maintaining relatively small change with respect to the original model.

### 2018-03-30 paper/predSummary.tex
length fit

To address these challenges, we introduce the summary view (see Fig.~\ref{fig:teaser}(a)), which consists of a treemap, a histogram, and a scatterplot, to summarize the prediction results of the $10k$ examples and provide the ability to drill down to individual examples for detailed analysis.
%treemap for prediction
As illustrated in Fig.~\ref{fig:summaryView}, we utilized a treemap (a) to encode the different combinations of the ground truth label and the predicted label. The green treemap blocks correspond to examples with correct predictions, whereas the orange blocks indicate failures. The size of the block encodes the number of examples belonging to each category.
%
By clicking on the treemap blocks, we can narrow down the selection by focusing on specific scenarios.
%
As we select the ``E/E'' (ground truth: E-Entailment / predicted label: E-Entailment) category in the treemap (see Fig.~\ref{fig:summaryView}(a)), the histogram (Fig.~\ref{fig:summaryView}(b)) and scatterplot  (Fig.~\ref{fig:summaryView}(c)) are shown. The histogram shows the distribution of the prediction sensitivity in the selected category. For each example, the stability is defined by the ratio of the number of perturbed pairs with the same predicted label and the all the perturbed pairs. Considering we generated 100 pairs via the automated sentence perturbation (i.e., replace nouns and verbs with synonymous), the stability is 0.8 if 80 out of the 100 maintain the original prediction label. 
%The \emph{stability} measures how likely a sentence pair will change its prediction after applying minor perturbations.% which helps the expert infer the model stability on the given example.
%
We can further narrow down the focused set by selecting the bins in the histogram (see Fig.~\ref{fig:summaryView}(b),(d)).
%
In the scatterplot, besides showing the sensitivity value, we also include the number of perturbed pairs (labeled as \emph{perturbCount}) to help users better assess the \emph{stability} number (i.e., if the \emph{perturbCount} is rather small ($<10$), then the \emph{sensitivity} number is likely to be quite noisy). %For example, if the \emph{perturbCount} is rather small ($<10$), then the \emph{sensitivity} number is likely to be quite noisy and less trustworthy, as compared to the sensitivity ratio computed from a much larger number of perturbed pairs.

>>>>

To address these challenges, we introduce the summary view (see Fig.~\ref{fig:teaser}(a)), which consists of a treemap, a histogram, and a scatterplot, to summarize the prediction results of the $10k$ examples and provide the ability to drill down to individual examples for detailed analysis.
%treemap for prediction
As illustrated in Fig.~\ref{fig:summaryView}, we utilized a treemap (a) to encode the different combinations of the ground truth label and the predicted label. The green treemap blocks correspond to examples with correct predictions, whereas the orange blocks indicate failures. The size of the block encodes the number of examples belonging to each category.
%
By clicking on the treemap blocks, we can narrow down the selection by focusing on specific scenarios.
%
As we select the ``E/E'' (ground truth: E-Entailment / predicted label: E-Entailment) category in the treemap (see Fig.~\ref{fig:summaryView}(a)), the histogram (Fig.~\ref{fig:summaryView}(b)) and scatterplot  (Fig.~\ref{fig:summaryView}(c)) are shown. The histogram shows the distribution of the prediction sensitivity in the selected category. For each example, the stability is defined by the ratio of the number of perturbed pairs with the same predicted label and the all the perturbed pairs. Considering we generated 100 pairs via the automated sentence perturbation (i.e., replace nouns and verbs with synonymous), the stability is 0.8 if 80 out of the 100 maintain the original prediction label. 
%The \emph{stability} measures how likely a sentence pair will change its prediction after applying minor perturbations.% which helps the expert infer the model stability on the given example.
%
We can further narrow down the focused set by selecting the bins in the histogram (see Fig.~\ref{fig:summaryView}(b),(d)).
%
In the scatterplot, besides showing the sensitivity value, we also include the number of perturbed pairs (labeled as \emph{perturbCount}) to help users better assess the \emph{stability} number (i.e., if the \emph{perturbCount} is rather small ($<10$), then the \emph{stability} number is likely to be quite noisy). %For example, if the \emph{perturbCount} is rather small ($<10$), then the \emph{sensitivity} number is likely to be quite noisy and less trustworthy, as compared to the sensitivity ratio computed from a much larger number of perturbed pairs.
